---
title: "SDS_315_HW_3"
author: "Neena Varanasi"
date: "2024-01-30"
output: 
  pdf_document
---
Name: Neena Varanasi

EID: nv6558

Github Link:

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message = FALSE, echo=FALSE}
library(tidyverse)
library(ggplot2)
library(kableExtra)
```

## Problem 1
```{r, message = FALSE, echo=FALSE}
#import dataset
creatinine <- read_csv("creatinine.csv")

#create linear model
model_kidney = lm(creatclear ~ age, data=creatinine)
kidney_values <- coef(model_kidney)


#extract variables and store linear equation
slope <- kidney_values[2]
intercept <- kidney_values[1]

```
A) The clearance rate for a 55-year old should be `r round(predict(model_kidney, newdata = data.frame(age = 55)),2)`. I determined this by creating a fitted model for the creatinine dataset and then using the predict function to determine the expected creatinine clearance rate for a 55-year-old. The fitted model reflects this equation: `r paste("y =", round(intercept, 2), "+", round(slope, 2), "* x")`.

B) Creatine clearance rate decreases with age as the rate of change is of the graph is`r round((slope),2)` I determined this value by creating a fitted linear model for the creatinine dataset. Then, I stored the coefficients of the graph (slope and intercept) in a list and extracted the slope from the list.

C) A 40-year-old with a rate of 135 is healthier as the 40-year-old's rate is higher than the average for their age by more than 10 mL/minute. The average clearance rate for a 40 year old is `r round(predict(model_kidney, newdata = data.frame(age = 40)),2)` Meanwhile, the 60-year-old's rate of 112 is  barely above the average rate for their age, which is `r round(predict(model_kidney, newdata = data.frame(age = 60)),2)`. The fitted model reflects this equation: `r paste("y =", round(intercept, 2), "+", round(slope, 2), "* x")`

## Problem 2
```{r, message = FALSE, echo=FALSE}
marketmodel <- read_csv("marketmodel.csv")

#linearly regress the returns of the stocks- s&p is the x-axis,
aapl <- lm(AAPL ~ SPY, data=marketmodel)
goog <- lm(GOOG ~ SPY, data=marketmodel)
mrk <- lm(MRK ~ SPY, data=marketmodel)
jnj <- lm(JNJ ~ SPY, data=marketmodel)
wmt <- lm(WMT ~ SPY, data=marketmodel)
tgt <- lm(TGT ~ SPY, data=marketmodel)

#extract the intercepts and slopes
intercepts <- c(coef(aapl)[1], coef(goog)[1], coef(mrk)[1], coef(jnj)[1], coef(wmt)[1], coef(tgt)[1])

slopes <-  c(coef(aapl)[2], coef(goog)[2], coef(mrk)[2], coef(jnj)[2], coef(wmt)[2], coef(tgt)[2])

r_squared <- c(summary(aapl)$r.squared, summary(goog)$r.squared, summary(mrk)$r.squared, summary(jnj)$r.squared, summary(wmt)$r.squared, summary(tgt)$r.squared)

#create a tibble
stock_table <- data.frame(
  TickerSymbol = c("AAPL", "GOOG", "MRK", "JNJ", "WMT", "TGT"),
  Slopes = slopes,
  Intercepts = intercepts,
  R_Squared = r_squared)

stock_table %>%
  kbl(digits = 2, col.names = c("Ticker Symbol", "Slopes", "Intercepts", "R^2 Value"), booktabs = TRUE) %>%
  kable_styling() %>% 
  kable_styling(latex_options = c("hold_position"))

```
The table above demonstrates the regression of the returns for 6 individual stocks, Apple (APPL), Google(GOOG) Merck (MRK), Johnson and Johnson (JNJ), Wal-Mart (WMT), and Target (TGT) with respect to the S&P 500. The ticker symbol, slope, intercepts, and R^2 values are shown for each of the respective stocks starting 2019.

The beta of a stock is a measure of the systematic risk of stocks that is represented similarly to percent change. A beta can have numerous values that correspond with different meanings. A beta of less than one means less systematic risk than the average firm. Meanwhile, a beta greater than one means the systematic risk is above average. This measure of risk means that when the market goes up, there is a greater return on the stock, but when the market goes down there are greater losses. A beta of 0 means no systematic risk, while a negative beta means that the return is less than the risk-free rate, similar to insurance. In the regression model above, the beta value with respect to the S&P 500 is represented by the slopes column. Therefore, Apple (AAPL) has an above-average risk, while Merck (MRK), Johnson and Johnson (JNJ), Wal-Mart (WMT), and Target (TGT) have below-average risks. An explanation for this below-average risk for Wal-Mart and Target could be that people buy necessities no matter the economic climate. Google (GOOG) represents the average firm with a systematic risk of 1.

Therefore, the firm with the lowest systematic risk is Wal-Mart, and the firm with the highest systematic risk is Apple.

## Problem 3
```{r, message = FALSE, echo=FALSE}
#import dataset
covid <- read_csv("covid.csv")

#subset of Italy data
italy_covid <- filter(covid, country == "Italy")

model_italy_covid <- lm(log(deaths) ~ days_since_first_death, data=italy_covid)
italy_values <- coef(model_italy_covid)

italy_growth_rate <- italy_values[2]
italy_doubling_time <- (70 / (italy_growth_rate *100))

#subset of Spain data
spain_covid <- filter(covid, country == "Spain")

model_spain_covid <- lm(log(deaths) ~ days_since_first_death, data=spain_covid)
spain_values <- coef(model_spain_covid)

spain_growth_rate <- spain_values[2]
spain_doubling_time <- (70 / (spain_growth_rate *100))

ggplot(covid,  aes(x=days_since_first_death, y=deaths, color = country)) +
  geom_line() + 
  ggtitle("Daily COVID-19 Deaths Over Time") +
  xlab("Days Since First Death") +
  ylab("Number of Deaths")

```
The growth rate for COVID-19 cases in Italy is `r round(italy_growth_rate,3)`. The doubling time for cases in Italy is `r round(italy_doubling_time)` days.

The growth rate for COVID-19 cases in Spain is `r round(spain_growth_rate,3)`. The doubling time for cases in Spain is `r round(spain_doubling_time)` days.

## Problem 4
```{r, message = FALSE, echo=FALSE}
#import dataset
milk <- read_csv("milk.csv")

#create linear model
milk_model <- lm(log(sales) ~ log(price), data=milk)	
milk_values <- coef(milk_model)

#extract elasticity
milk_price_elasticity <- milk_values[2]
```
The price elasticity of milk is `r round(milk_price_elasticity,2)`. In order to get this estimate, I used the power law as the relationship between sales and price is relatively proportional. Therefore, I fitted the model using the natural logarithm of each variable and extracted the coefficient on x, which represents the elasticity.

